---
  title: "Functional Programming Concepts"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Closure Functions

Closure functions are functions written by functions. Closures get their name because they enclose the environment of the parent function and can access all its variables. This is useful because it allows us to have two levels of parameters: a parent level that controls operation and a child level that does the work

Use the mtcars data frame as an example.
```{r load_data}
attach(mtcars)
mtcars$model <- rownames(mtcars)
```

A closure function that calculates the power to weight ratio. The below horsepwr_to_kw_conv and lbs_to_kg functions are also examples of Function Factory patterns.
```{r func1}
# genertic reusable convert function
convert_fun <- function(conv_rate) {
  function(x) {
    x*conv_rate
  }
}
# horse power to kw unit conversion
horsepwr_to_kw_conv <- 0.7457
# horse power to kw function
horsepwr_to_kw <- convert_fun(horsepwr_to_kw_conv)
# pounds to kg unit conversion
lb_to_kg_conv <- 0.453592
# lbs to kg function
lbs_to_kg <- convert_fun(lb_to_kg_conv)
# power (kw) to weight (kg) function
power_to_weight_func <- function(wt_in) {
  function(hp_in) {
    horsepwr_to_kw(hp_in)/(lbs_to_kg(wt_in))  
  }
}
power_to_weight_kg <- power_to_weight_func(mtcars$wt)
power_to_weight_kg(mtcars$hp)
```