---
  title: "Functional Programming Concepts"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monoids

When you deal with various types in a programming language you will notice common patterns of behavior for binary operations. Here are some examples:

* Integers with addition. We know that (a+b)+c == a+(b+c) and 0+n == n+0 == n; same with multiplication: (a*b)*c == a*(b*c) and 1*n == n*1 == n
* Strings with concatenation. a+(b+c)==(a+b)+c; ""+s==s and s+"" == s, etc.
* Lists with concatenation, like List(1,2)+List(3,4) == List(1,2,3,4).
* Sets with their union, like Set(1,2,3)+Set(2,4) == Set(1,2,3,4).

A common pattern with a data type, a binary operation, and a special instance of the type, and certain rules is called monoid.

Below is an example showing a monoid pattern to calculate the sum of multiple vectors.
```{r func1}
a = c(rep(1:10))
b = c(rep(11:20))
c = c(rep(21:30))
all = list(a, b, c)
add_func <- function() {
  y <- 0
  function(x) {
    y<<-y + sum(x)
    y
  }
}

add_func1 <- add_func()

z = lapply(all, add_func1)

```