---
  title: "Functional Programming Concepts"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Partial Functions
Partial function application allows you to modify a function by pre-filling some of the arguments. It is particularly useful in conjunction with functionals and other function operators.

Use the mtcars data frame as an example.
```{r load_data}
library(pryr)
attach(mtcars)
mtcars$model <- rownames(mtcars)
```

A closure function that calculates the power to weight ratio. The below horsepwr_to_kw_conv and lbs_to_kg functions are also examples of Function Factory patterns.
```{r func1}
# genertic reusable convert function
convert_fun <- function(conv_rate) {
  function(x) {
    x*conv_rate
  }
}
# horse power to kw unit conversion
horsepwr_to_kw_conv <- 0.7457
# horse power to kw function
horsepwr_to_kw <- convert_fun(horsepwr_to_kw_conv)
# pounds to kg unit conversion
lb_to_kg_conv <- 0.453592
# lbs to kg function
lbs_to_kg <- convert_fun(lb_to_kg_conv)
# power (kw) to weight (kg) function
power_to_weight_func <- function(hp_in, wt_in) {
  horsepwr_to_kw(hp_in)/(lbs_to_kg(wt_in))  
}
power_to_weight_func_new = pryr::partial(power_to_weight_func, wt_in=mtcars$wt)
power_to_weight_func_new(hp_in=mtcars$hp)
```