---
title: "ISIL_2_8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ISIL_2_8

This is a short excerpt of the tutorials from the ISIL book (Chapter 2, Question 8) 

## Required Libraries
Load the required librarys for the question
```{r library, include=FALSE}
library(ggplot2)
library(reshape2)
library(PerformanceAnalytics)
```

## Data
  Load the data from the csv file
```{r college}
college = read.csv("College.csv", header = TRUE, sep = ",")
```

  Clean the data of the university name column, and add them as row ids
```{r clean_college, echo=FALSE, include=FALSE}
#Change the numerical row numbers to the university row name
rownames(college) = college[,1]
#fix(college)
#Remove the university column
college = college[,-1]
#8(c)
# i Obtain a summary of the data frame
```
  Create a new feature called Elite, based on if there are more than 50 students from the top 10% of high school graduates
```{r create_Elite_feature, echo=FALSE}
college$Elite = "No"
college[ which(college$Top10perc>50), ]$Elite = "Yes"
college$Elite = as.factor(college$Elite)
```

Below is a summary of the college data:
```{r summary_college}
summary(college)
```

## Visualisations

##Elite vs Non-Elite
We want to find out if there are any distribution differences between elite and non-elite universities:

```{r plot_elite_vs_nonelite, echo=FALSE, warning=FALSE, fig.width=20, fig.height=20}
quant = melt(college, id=as.array(names(Filter(is.factor, college))))
ggplot(quant, aes(x = value, group=Elite, fill=Elite)) + 
  facet_wrap(~variable,scales = "free") + 
  geom_histogram(aes(y = ..density..), position="dodge", binwidth = 0.01) +
  geom_density(alpha=0.4)+
  scale_x_log10()+
  labs(title="Elite histogram plot",x="Count", y = "Density")+
  theme_minimal() +
  theme(strip.text = element_text(size = rel(2)))+
  scale_color_brewer(palette = "Paired")
cat("  \n")
```

## Inferential Analysis - Elite vs. Non-Elite
```{r t_test, echo=FALSE}
nums_col <- sapply(college, is.numeric)
t(sapply(college[,nums_col], function(x) 
     unlist(t.test(x~college$Elite)[c("estimate","p.value","statistic","conf.int")])))
```
* The Elite universities receive, enrol, and accept more applications than the other universities
* They have more full-time, but fewer part-time undergraduates than the other universities
* They receive more out of state tuition than the other universities. It has a negatively skewed distribution. Why is this so? (futher analysis below)
```{r elite_and_non_elite_outstate, echo=FALSE}
elite_outstate = college[which(college$Elite=='Yes'),]$Outstate
non_elite_outstate = college[which(college$Elite=='No'),]$Outstate
```
+ As shown in the below boxplot, the Elite universities have a slightly larger IQR, and a lower mean than median value, which is reflected in its negative skew. The Non-elite university has a slightly more normal distribution:


```{r plot_boxplot_out_state, echo=FALSE, warning=FALSE, fig.width=3, fig.height=3}
ggplot(college, aes(x = Elite, y=Outstate)) + 
  geom_boxplot() +
  theme_minimal() +
  theme(strip.text = element_text(size = rel(2)))
cat("  \n")
```

Stat    | Elite                                                             |  Non-Elite
------- | -------------------------------------------                       | ---------------------------------------------
Mean    | `r format(mean(elite_outstate), digits = 7)`                      | `r format(mean(non_elite_outstate), digits = 7)`
Median  | `r format(median(elite_outstate), digits = 7)`                    | `r format(median(non_elite_outstate), digits = 7)`
25th    | `r format(quantile(elite_outstate, probs=c(0.25)), digits = 7)`   | `r format(quantile(non_elite_outstate, probs=c(0.25)), digits = 7)`
75th    | `r format(quantile(elite_outstate, probs=c(0.75)), digits = 7)`   | `r format(quantile(non_elite_outstate, probs=c(0.75)), digits = 7)`


* The room board is more expensive than the other universities, with a bimodal distribution (a dip in price in the middle of the distribution). What is causing this? Is it due to varying real estate prices based on state? (futher analysis below)

```{r plot_elite_vs_nonelite_roomboard, echo=FALSE, warning=FALSE, fig.width=5, fig.height=5}
ggplot(quant[which(quant$variable =='Room.Board' & quant$value>3000 & quant$value<7000),], aes(x = value, group=Elite, fill=Elite)) +   geom_histogram(aes(y = ..density..), position="dodge", binwidth = 100) +
  geom_density(alpha=0.4)+
  labs(title="Elite histogram plot",x="Count", y = "Density")+
  theme_minimal() +
  theme(strip.text = element_text(size = rel(2)))+
  scale_color_brewer(palette = "Paired")
cat("  \n")
```
  + The dip occurs between 4500 - 5500. As shown below, the dip coincides with a lower Expenditure.
```{r, echo=FALSE}
college_elite_tmp = college[which(college$Elite=="Yes"),]
college_elite_tmp$dip_room_board = "No"
college_elite_tmp[which(college_elite_tmp$Room.Board>4500 & college_elite_tmp$Room.Board<5500),]$dip_room_board = "Yes"
college_elite_tmp$dip_room_board = as.factor(college_elite_tmp$dip_room_board)
quant = melt(college_elite_tmp, id=as.array(names(Filter(is.factor, college_elite_tmp))))
```
```{r plot_dip_in_roomw_board, echo=FALSE, warning=FALSE, fig.width=20, fig.height=20}
ggplot(quant, aes(x = value, group=dip_room_board, fill=dip_room_board)) + 
  facet_wrap(~variable,scales = "free") + 
  geom_histogram(aes(y = ..density..), position="dodge", binwidth = 0.1) +
  geom_density(alpha=0.4)+
  scale_x_log10()+
  labs(title="Private histogram plot",x="Count", y = "Density")+
  theme_minimal() +
  theme(strip.text = element_text(size = rel(2)))+
  scale_color_brewer(palette = "Paired")
cat("  \n")
```
```{r, echo=FALSE}
nums_col <- sapply(college_elite_tmp, is.numeric)
t(sapply(college_elite_tmp[,nums_col], function(x) 
     unlist(t.test(x~college_elite_tmp$dip_room_board)[c("estimate","p.value","statistic","conf.int")])))
cat("  \n")
```


* The personal spending are approximately similar around the mean, but there are fewer Elite university students that tend to spend less after the mean. (further analysis below)
* There are higher number of PhDs, and Teriminal degrees at Elite universities
* There is a lower Student to faculty ratio at Elite universitites. Is it because there are more faculties at Elite universities or more students at Non-elite universities?
* The Instructional Expenditure per Student, Percentage of Alumni who Donate, and Graduation Rates are higher at Elite universities.


##Private vs Non-Private
As well as Private vs. Non-Private:


```{r plot_private_vs_nonprivate, echo=FALSE, warning=FALSE, fig.width=20, fig.height=20}
ggplot(quant, aes(x = value, group=Private, fill=Private)) + 
  facet_wrap(~variable,scales = "free") + 
  geom_histogram(aes(y = ..density..), position="dodge", binwidth = 0.01) +
  geom_density(alpha=0.4)+
  scale_x_log10()+
  labs(title="Private histogram plot",x="Count", y = "Density")+
  theme_minimal() +
  theme(strip.text = element_text(size = rel(2)))+
  scale_color_brewer(palette = "Paired")
cat("  \n")

```

## Inferential Analysis - Private vs. Non-Private

```{r t_test_private, echo=FALSE}
nums_col <- sapply(college, is.numeric)
t(sapply(college[,nums_col], function(x) 
     unlist(t.test(x~college$Private)[c("estimate","p.value","statistic","conf.int")])))
```
