---
  title: "Functional Programming Concepts"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mutable States

Having variables at two levels allows you to maintain state across function invocations. This is possible because while the execution environment is refreshed every time, the enclosing environment is constant. The key to managing variables at different levels is the double arrow assignment operator (<<-). Unlike the usual single arrow assignment (<-) that always assigns in the current environment, the double arrow operator will keep looking up the chain of parent environments until it finds a matching name. (Binding names to values has more details on how it works.)

Together, a static parent environment and <<- make it possible to maintain state across function calls. The following example shows a counter that records how many times a function has been called. Each time new_counter is run, it creates an environment, initialises the counter i in this environment, and then creates a new function.

```{r func1}
counter_func <- function(increment_val=1) {
  count_val <- 0
  function() {
    count_val <<- count_val + increment_val
    count_val
  }
}

counter_func1 <- counter_func()
counter_func2 <- counter_func(2)
```

Below shows an example of the states not being maintained. 
```{r func2}
xcounter_func <- function(increment_val=1) {
  count_val <- 0
  function() {
    count_val <- count_val + increment_val
    count_val
  }
}

xcounter_func1 <- xcounter_func()
xcounter_func2 <- xcounter_func(2)
```

