---
title: "ISIL_6_5"
output: github_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ISIL_6_5

### 6.5.1 Best Subset

The regsubsets() function (part of the leaps library) performs best subregsubsets() set selection by identifying the best model that contains a given number of predictors, where best is quantified using RSS. The syntax is the same as for lm()

Load the required librarys for the question
```{r library, include=FALSE}
library(ISLR)

predict_regsubsets<- function(object, newdata, id, ...) {
  form = as.formula(object$call[[2]])
  mat = model.matrix(form, newdata)
  coefi = coef(object, id=id)
  xvars = names(coefi)
  mat[, xvars] %*% coefi
}

```

Load the data
```{r load_data}
fix(Hitters)
names(Hitters)
dim(Hitters)
sum(is.na(Hitters))

```

Omit players without salaries
```{r omit_na}
Hitters <- na.omit(Hitters)
dim(Hitters)
sum(is.na(Hitters))
```

The regsubsets() function (part of the leaps library) performs best subregsubsets() set selection by identifying the best model that contains a given number of predictors, where best is quantified using RSS.

```{r reg_subset}
library(leaps)
regfit_Hitters <- regsubsets(Salary~., Hitters, nvmax = 19)
reg_summary =summary (regfit_Hitters)
reg_summary
```

An asterisk indicates that a given variable is included in the corresponding model. For instance, this output indicates that the best two-variable model contains only Hits and CRBI.

Plot the R square.
```{r plot_rsq_1}
names(reg_summary)
par(mfrow =c(2,2))
plot(reg_summary$rss , xlab="Number of Variables", ylab="RSS", type="l")
plot(reg_summary$adjr2 , xlab="Number of Variables", ylab="RSq", type="l")
r2_max <- which.max(reg_summary$adjr2)
points(r2_max, reg_summary$adjr2[r2_max], col="red", cex=2, pch=20)
```

Plot the BIC.
```{r plot_cp_bic_1}
plot(reg_summary$cp , xlab="Number of Variables", ylab="Cp", type="l")
best <- which.min(reg_summary$cp)
points(best, reg_summary$cp[best], col="red", cex=2, pch=20)
plot(reg_summary$bic , xlab="Number of Variables", ylab="BIC", type="l")
best <- which.min(reg_summary$bic)
points(best, reg_summary$bic[best], col="red", cex=2, pch=20)
```

Plot all.
```{r plot_regsubset_all}
plot(regfit_Hitters, scale='r2')
plot(regfit_Hitters, scale='adjr2')
plot(regfit_Hitters, scale='Cp')
plot(regfit_Hitters, scale='bic')
```

The coeffecient estimates
```{r regsubset_coeffestimates}
coef(regfit_Hitters, 6)
```

### 6.5.2 Stepwise Selection

Forward stepwise
```{r fwd_stepwise}
regfit_fwd <- regsubsets(Salary~., data=Hitters, nvmax=19, method="forward")
summary(regfit_fwd)
```

Backward stepwise
```{r bwd_stepwise}
regfit_bwd <- regsubsets(Salary~., data=Hitters, nvmax=19, method="backward")
summary(regfit_bwd)
```

Compare different subsetting selected predictors
```{r subset_compare}
coef(regfit_Hitters, 7)
coef(regfit_fwd, 7)
coef(regfit_bwd, 7)
```

### 6.5.3 Cross validation
```{r cv_sample}
set.seed(1)
train <- sample(c(TRUE, FALSE), nrow(Hitters), rep=TRUE)
test <- (!train)
regfit_best <- regsubsets(Salary~., data=Hitters[train, ], nvmax=19)
test_matrix <- model.matrix(Salary~., data=Hitters[test, ])
val_errors <- rep(NA, 19)
for (i in 1:19) {
  coefi <- coef(regfit_best, id=i)
  # %*% is a matrix multiplication operator
  pred = test_matrix[, names(coefi)] %*% coefi
  val_errors[i] = mean((Hitters$Salary[test] - pred)^2)
}
```

Best performing model has 10 predictors
```{r cv_best_model1}
val_errors
which.min (val_errors)
coef(regfit_best, 10)
```

Choose models of different sizes using CV, by executing best subset selection within each k training sets.
```{r cv_best_model2}
k <- 10
set.seed(1)
folds <- sample(1:k, nrow(Hitters), replace = TRUE)
cv_errors <- matrix(NA, k, 19, dimnames=list(NULL, paste(1:19)))
for (j in 1:k) {
  best_fit <- regsubsets(Salary~., data=Hitters[folds !=j, ], nvmax=19)
  for(i in 1:19) {
    pred <- predict_regsubsets(best_fit, Hitters[folds==j, ], id=i)
    cv_errors[j, i] <- mean((Hitters$Salary[folds==j] - pred)^2)
  }
}
mean_cv_errors <- apply(cv_errors, 2, mean)
mean_cv_errors
par(mfrow=c(1, 1))
plot(mean_cv_errors, type='b')
```


